---
title: "SQL Analysis on Stack Exchange"
author: "Wanzhu Zheng"
date: "2023"
output:
  pdf_document:
    latex_engine: xelatex
  'pdf_document: latex_engine: xelatex': default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminary Setup

```{r}
library(DBI)
library(RSQLite)
db = dbConnect(RSQLite::SQLite(), "/Users/wanzhu_zheng/Downloads/stats.stackexchange.db")
dbListTables(db)
lapply(dbListTables(db), function(tbl) dbListFields(db, tbl))
```

Q1: How many users are there?  
This question asks for the total number of users that are signed up on the
StatsExchange Website. Users in this case refers to the Id in the Users table 
```{r}
users =  dbGetQuery(db, "SELECT Id,
                    COUNT(DISTINCT Id) AS Num 
                    FROM Users GROUP BY Id")
head(users)
sum(users$Num)
```
Id column refers to the distinct user Id's and Num refers to the number of
users with that Id. As expected, every user Id has a value of 1.
From the results, we see there are 321677 total users.  


Q2: How many users joined since 2020?  
This question asks for the total number of users that joined since 2020.
```{r}
new_users = dbGetQuery(db, "SELECT CreationDate, 
                       COUNT(DISTINCT Id) AS Num FROM Users 
                       GROUP BY CreationDate 
                       HAVING strftime(CreationDate) > '2020'")
head(new_users)
sum(new_users$Num)
```
We create a dataframe that shows when an account was created (CreationDate)
and how many accounts were created at that time (Num). From the results,
there are 100796 that joined since 2020.  

Q3: How many users joined each year?  
Describe this with a plot, commenting on any anomalies.  
We want to find out how many users join the forum every year.  
```{r}
new_users_year = dbGetQuery(db, "SELECT SUBSTRING(CreationDate, 1, 4) AS Joined, 
                            COUNT(*) AS Num 
                            FROM Users 
                            GROUP BY Joined")
new_users_year
plot(Num ~ Joined, data=new_users_year, main = "Number of New Users per Year", 
     xlab="Year Joined", ylab="Number of New Users")
axis(side=1, at=seq(2010, 2023, by=1))
axis(side=2, at=seq(0, 45000, by=5000))
```
We return a dataframe that shows the year each user joined from 2010, and the
total number of new users each year. We plot this in a scatterplot which shows
that the number of new users peaked in 2017 with 44416 users joining.  

Q4: How many different types of posts are there in the Posts table? 
Get the description of the types from the PostTypeIdMap table.     
In other words, create a table with the description of each post type and the 
number of posts of that type, and arrange it from most to least occurrences.
We want to see how many different types of posts a user can post. We will
extract this from the Posts table.  
```{r}
type_query = "
SELECT * 
FROM (
  SELECT Id, value AS Description
  FROM PostTypeIdMap
) AS description
LEFT JOIN (
  SELECT PostTypeId AS Type, COUNT(*) AS Num
  FROM Posts
  GROUP BY PostTypeId
  ORDER BY Num DESC
) AS types
ON description.Id = types.Type"

type_freq = dbGetQuery(db, type_query)
type_freq$Type = NULL
type_freq = type_freq[order(type_freq$Num, decreasing=TRUE), ]
type_freq
```
To do so, we create a dataframe with three columns. Id describes the Id
associated with each post type, Description is the description of the post
for each post Id, and Num describes the number of posts that fall under
each category.  

Q5: How many posted questions are there?  
We want to find the total number of questions posted on the forum.
To do so, we need a conditional statement that specifies the post type Id
as 1, as seen from our previous answer.  
```{r}
type_freq$Num[type_freq$Id == 1]
```
To find the answer, we just need to call the type_freq table from Q4 and
specify that the Id is 1. There are 204370 questions.  

Q6: What are the top 50 most common tags on questions? For each of the top 50 
tags on questions, how many questions are there for each tag?  
We want to find the top 50 most common tags found on all the questions. Then, 
we want to find how many questions are associated with each of the top 50 tags.
```{r}
tags_50 =  dbGetQuery(db, "SELECT Tag, COUNT(*) AS Num 
                   FROM TagPosts 
                   GROUP BY Tag
                   ORDER BY Num DESC
                   LIMIT 50")
tags_50
```
We return a dataframe with the column Tag that has all the tag names, and
order the frequency the tags show up by Num and only return the top 50
greatest numbers.  

Q7: How many tags are in most questions?   
This question asks how many tags are present in most questions.  
```{r}
tags_per_q =  dbGetQuery(db, "SELECT Id, COUNT(Tag) AS Num 
                         FROM TagPosts GROUP BY Id 
                         ORDER BY Num ASC")
head(tags_per_q)
tail(tags_per_q)
summary(tags_per_q$Num)
```
We first return a dataframe that shows the number of tags per question Id.
This tells us that the number of tags ranges anywhere between 1 to 5.
Then, we call the summary function which shows that there's a median of 3
tags, and a mean of 3.085 tags.  

Q8: How many answers are there?   
Similar to Q5, this question is asking us to identify the total number of 
answers that are posted.  
```{r}
type_freq$Num[type_freq$Id == 2]
```
There are 197928 answers posted.  

Q9: What’s the most recent question (by date-time) in the Posts table?  
This question asks us to convert the dates to datetime formate and use that
to return the most recent question that was posted on the forum.
```{r}
creation_query = "SELECT MAX(strftime(CreationDate)) AS Date, Body AS Question
                         FROM Posts WHERE PostTypeId='1'"
newest_post = dbGetQuery(db, creation_query)
newest_post$Date
newest_post$Question
```
The most recent question was posted on 03-05-2023 and asks if there
are any methods to combine VI and MCMC. The URL is https://stats.stackexchange.com/questions/608458/are-there-any-methods-that-combine-mcmc-and-vi.
If there was a URL field in Posts, we would map it by calling URL from Posts
so our statement would look like "SELECT MAX(strftime(CreationDate)) AS Date, 
Body AS Question, URL
FROM Posts WHERE PostTypeId='1'"  

Q14: What is the date range for the questions and answers in this database?  
I interpreted the date range as finding the first question/answer and the
most recent question/answer.
```{r}
date_range_q = dbGetQuery(db, "SELECT MIN(strftime(CreationDate)) AS Min_Date,
                        MAX(strftime(CreationDate)) AS Max_Date FROM Posts
                        WHERE PostTypeId='1'")
date_range_q
date_range_a = dbGetQuery(db, "SELECT MIN(strftime(CreationDate)) AS Min_Date,
                        MAX(strftime(CreationDate)) AS Max_Date FROM Posts
                        WHERE PostTypeId='2'")
date_range_a
```

The first question was posted on 02-02-2009 and the most recent question was
posted on 03-05-2023. The first answer was posted on 02-02-2009 and the most
recent question was posted on 03-05-2023. This is expected since the first
and most recent questions posted got their response on the same day, 
but at a later time.  

Q15: What question has the most comments associated with it? 
How many answers are there for this question?  
This question asks to return the question with the most comments and asks
us to find the number of answers to the question
```{r}
most_comments = dbGetQuery(db, "SELECT Id, Body, CommentCount, 
                           AnswerCount FROM Posts
                           WHERE PostTypeId = '1'
                           ORDER BY CommentCount DESC
                           LIMIT 1")
most_comments
```
We return a dataframe with the question Id, the question itself, how many
comments are associated with it (54), and how many answers the question has (6).  

Q16: How many comments are there across all posts? 
How many posts have a comment?
What is the distribution of comments per question?  
This asks for the total number of comments for all types of posts. Then asks
to find how many posts have a comment. Finally, we return the distribution
fo comments per question in a histogram.
```{r}
comments = dbGetQuery(db, "SELECT Id, CommentCount FROM POSTS
                      ORDER BY CommentCount DESC")
head(comments)
sum(comments$CommentCount)

has_comment = dbGetQuery(db, "SELECT Id, CommentCount FROM POSTS
                         WHERE CommentCount != 0
                         ORDER BY CommentCount DESC")
nrow(has_comment)

comments_per_q = dbGetQuery(db, "SELECT Id, CommentCount FROM POSTS
                            WHERE PostTypeId='1'
                            ORDER BY CommentCount DESC")
head(comments_per_q)
comments_freq = table(comments_per_q$CommentCount)
comments_freq

hist(comments_per_q$CommentCount,breaks=seq(0,70,l=54),
     xlab="Number of Comments", 
     main="Distribution of Comments over Questions")
```
We first create a dataframe that shows the number of comments for each Id that
is returned from most comments to least comments. Then, we add up all the
comments to find that there are 768069 comments in total. Then we create another
dataframe where our condition statement is all posts that have comments. There
are 229859 posts that have comments. We then return a histogram mapping the
frequency of comments. For eample, we see that the histogram is right-skewed,
meaning that most posts have 0 comments, as seen by our frequency table.  

Q18: Do the people who vote tend to have badges?
This asks us to find the relationship between people who vote and people who
have badges.
```{r}
voters = dbGetQuery(db, "SELECT COUNT(DISTINCT Votes.UserId) AS voters FROM VOTES")
voters

votes_query = "
SELECT COUNT(DISTINCT Votes.UserId) AS voters
FROM VOTES
JOIN Badges ON Votes.UserId = Badges.UserId
"
voters_badges = dbGetQuery(db, votes_query)
voters_badges
```
We first return a dataframe that tells us how many voters on on the forum.
From our results we see that there are 3191 voters in total. Then we return
a dataframe that shows us how many people voters have badges by joining the 
two query tables. We see that 3186 of voters have badges. This suggests that
most voters tend to have badges.  

Q21: Compute the table that contains the question, the name of the user who
posted it, when the user joined, their location, the date the question was
first posted, the accepted answer, when the accepted answer was posted, 
the name of the user who provided the accepted answer.
```{r}
master_query = "
SELECT 
Q.Body AS Question, A.Body AS AcceptedAnswer, Q.Id AS QuestionId, 
A.Id AS AcceptedAnswerId, Q_User.DisplayName AS PosterUsername, 
Q_User.CreationDate AS PosterJoined, Q_User.Location AS Location,
A_User.DisplayName AS AnswererUsername, Q.CreationDate AS DateFirstPosted, 
A.CreationDate AS DateFirstAnswered
FROM Posts AS Q, Posts AS A, Users AS Q_User, Users AS A_User
WHERE Q.OwnerUserId = Q_User.Id
AND A.OwnerUserId = A_User.Id
AND A.Id = Q.AcceptedAnswerId
ORDER BY Q.Id ASC"

master_table = dbGetQuery(db, master_query)
master_table$AcceptedAnswerId = NULL
master_table$QuestionId = NULL
head(master_table)
```

Q22: Determine the users that have only posted questions and never 
answered a question? How many are there? 
This questions asks us to compute the number of questions, number of answers, 
and the user’s login name for this group. Then, we find users that have 0
answers.
```{r}
QA_query = "
SELECT
COUNT(DISTINCT CASE WHEN Q.PostTypeId = 1 THEN Q.Id END) AS QuestionCount, 
COUNT(DISTINCT CASE WHEN A.PostTypeId = 2 THEN A.Id END) AS AnswerCount, 
User.DisplayName AS UserName
FROM Posts AS Q, Posts AS A, Users AS User
WHERE Q.OwnerUserId = User.Id
AND A.OwnerUserId = User.Id
GROUP BY UserName
ORDER BY QuestionCount DESC
"

QA_table = dbGetQuery(db, QA_query)
only_questions = QA_table[QA_table$AnswerCount == 0, ]
head(only_questions)
nrow(only_questions)
```
From the datatable we see the username of the user, the number of questions
they posted, and the number of questions they answered, which should be 0. 
There are 63377 users that only asked questions and never answered quesitons.  

Q23: Compute the table with information for the 75 users with the most 
accepted answers. This table should include the user's display name, 
creation date, location, the number of badges they have won, the names of the
badges, the dates of the earliest and most recent accepted answer, and
the (unique) tags for all the questions for which they had the 
accepted answer (as a single string).
```{r}
accepted_query = "
SELECT
Users.DisplayName, Users.CreationDate AS Joined, Users.Location, 
COUNT(AA.AcceptedAnswerId) AS AcceptedCount, COUNT(Badges.Id) AS BadgeCount,
GROUP_CONCAT(DISTINCT Badges.Name) AS Badges, MIN(AA.CreationDate) AS EarliestAnswer,
MAX(AA.CreationDate) AS MostRecentAA, GROUP_CONCAT(DISTINCT Joiner.Tags) AS BadgeNames
FROM Users, Badges, Posts AS Joiner, Posts AS AA
WHERE Users.Id = Badges.UserId
AND AA.OwnerUserId = Users.Id
AND AA.Id = Joiner.AcceptedAnswerId
GROUP BY Users.DisplayName
ORDER BY COUNT(AA.AcceptedAnswerId) DESC
LIMIT 75
"

accepted_table = dbGetQuery(db, accepted_query)
head(accepted_table)
```
This table shows the user's display name, account creation date, 
location, the number of badges they have won, the names of the
badges, the dates of the earliest and most recent accepted answer, and
the (unique) tags for all the questions for which they had the 
accepted answer (as a single string).  

Q24: How many questions received no answers (accepted or unaccepted)? 
How many questions had no accepted answer?  
This questions asks us to find questions that recieved no answers in general
as well as no accepted answers.
```{r}
total_answers_query = "
SELECT
Q.Id AS QuestionId, Q.Body AS Question, COUNT(A.Id) AS AcceptedAnswerCount,
Q.AnswerCount AS TotalAnswerCount
FROM Posts Q
LEFT JOIN Posts A ON Q.AcceptedAnswerId = A.Id AND A.PostTypeId = 2
WHERE Q.PostTypeId = 1
GROUP BY Q.Id, Q.Body
HAVING AcceptedAnswerCount = 0 OR TotalAnswerCount = 0
ORDER BY TotalAnswerCount DESC
"

total_answers = dbGetQuery(db, total_answers_query)
head(total_answers)
no_answers = sum(total_answers$TotalAnswerCount == 0)
no_answers
no_accepted_answers = sum(total_answers$AcceptedAnswerCount == 0)
no_accepted_answers
```
We return a dataframe with the question Id, The question, the number of 
accepted answers, and the number of total answers. There are 66970 questions
with no answers and 136366 questions with no accepted answers.  

Q25: What is the distribution of answers per posted question?
We want to find the distribution of answers for each question.
```{r}
answers = dbGetQuery(db, "SELECT Id, AnswerCount FROM POSTS
                     WHERE PostTypeId = '1'
                     ORDER BY AnswerCount DESC")
head(answers)

answers_freq = table(answers$AnswerCount)
answers_freq

summary(answers$AnswerCount)

hist(answers$AnswerCount,breaks=seq(0,155,l=154),
     xlab="Number of Answers",
     main="Distribution of Answers Per Question")
```
We return a histogram that shows the frequency of how many answers each
question receives. We see that the histogram is skewed to the right with 
majority of answers being at 0. This matches our frequency table.  

Q26: What is the length of time for a question to receive an answer? 
To obtaining an accepted answer?  
This question asks for how long it takes for a question to receive an answer
and how long it takes to receive an accepted answer.
```{r}
answer_response_query = "
SELECT
Q.Id AS QuestionId,
(unixepoch(A.CreationDate) - unixepoch(Q.CreationDate)) / 60 AS AnswerTime
FROM Posts Q
JOIN Posts A ON Q.Id = A.ParentId
GROUP BY Q.Id
ORDER BY Q.Id ASC
"

acceptedanswer_response_query = "
SELECT
Q.Id AS QuestionId,
(unixepoch(AA.CreationDate) - unixepoch(Q.CreationDate)) / 60 AS AnswerTime
FROM Posts Q
JOIN Posts AA ON Q.AcceptedAnswerId = AA.Id
GROUP BY Q.Id
ORDER BY Q.Id ASC
"
a_response_length = dbGetQuery(db, answer_response_query)
head(a_response_length)

aa_response_length = dbGetQuery(db, acceptedanswer_response_query)
head(aa_response_length)
```
We return two dataframes. The first one tells us the response length in
minutes for how long it takes to receive an answer for each question. The 
second table tells us the response length it takes to receive an accepted
answer. These tables are orderd by question Id.  

Q27: How many answers are typically received before the accepted answer?
This questions asks us to find how many answers a question receives before
receiving the accepted answer.
```{r}
before_accepted_query = "
SELECT
Q.Id AS QuestionId,
(unixepoch(AA.CreationDate) - unixepoch(Q.CreationDate)) AS AATimeDifference,
(unixepoch(A.CreationDate) - unixepoch(Q.CreationDate)) AS ATimeDifference,
COUNT(A.Id) AS BeforeAccepted
FROM Posts AS Q, Posts AS AA, Posts AS A
WHERE Q.Id = A.ParentId
AND Q.AcceptedAnswerId = AA.Id
AND AATimeDifference < ATimeDifference
GROUP BY Q.Id
ORDER BY BeforeAccepted DESC
"

before_accepted = dbGetQuery(db, before_accepted_query)
head(before_accepted)
summary(before_accepted$BeforeAccepted)
```
We return a dataframe that shows the question Id, the time it takes for a 
question to receive an accepted answer (AATimeDifference), the time it takes 
for a question to receive an just an answer (ATimeDifference), and how many
answers were received before the accepted answer was posted (BeforeAccepted).
From the table, we can generalize that there was a median of 1 answer received
before the accepted answer and there was an average of 1.547 answers received
before the accepted answer.





